---
title: "Spatial Experiment"
output:
  html_document:
    df_print: paged
---

Spatial Transcriptomics based on https://lmweber.org/OSTA-book/analysis-steps.html 

For both practice data from github and from mice kidney paper

```{r, include=FALSE} 
library(SpatialExperiment)
library(STexampleData)
library(ggspavis)
library(scater)
library(scran)
library(pheatmap)

#load object from sample data
# spe <- Visium_humanDLPFC()
# imgData(spe)
```


Load the data and create a Spatial Experiment object
```{r, message=FALSE}

dir <- file.path("~","Desktop","test")

sample1 <-  file.path(dir, "week", "outs")

spe1 <- read10xVisium(sample1,
                        type = "sparse", data = "filtered",
                        images = "lowres", load = FALSE)


# sample_ids2 <- c("2day", '12hour')
# samples2 <- file.path(dir, sample_ids2, "outs")
# 
# samples_idsAll <- c('2day', '12hour', 'week', 'sham')
# samplesAll <-  file.path(dir, samples_idsAll, 'outs')
# 
# #Import visium data for mutiple samples
# spe2 <- read10xVisium(samples2, sample_ids2,
#                         type = "sparse", data = "filtered",
#                         images = "lowres", load = FALSE)
# 
# speAll <- read10xVisium(samplesAll, samples_idsAll,
#                         type = "sparse", data = "filtered",
#                         images = "lowres", load = FALSE)

```


```{r}

plotSpots(spe1)
# plotSpots(spe2)
# plotSpots(speAll)
# plotVisium(speAll)
plotVisium(spe1)
```

```{r}

# colData(spe)$sum <- colSums(counts(spe)) # add some values in 'colData' to annotate spots, also done by addPerCellQC function
# 
# # spe1 <- spe1[, colData(spe1)$in_tissue == TRUE] #All spots already in tissue because using filtered data
# 
# #rowData(spe1)$symbol #Get a list of gene names using rowData, symbol columns is the names
# 
# 

is_mito <- grepl("(^MT-)|(^mt-)", rowData(spe1)$symbol) #Used to look for specific genes, in examples searches for mitochondrial genes

# table(is_mito)
rowData(spe1)$symbol[is_mito]

spe1 <- addPerCellQC(spe1, subsets = list(mito = is_mito))
head(colData(spe1))

plotVisium(spe1, fill = "sum", trans = "log", highlight = "in_tissue")


```
Thresholds can be used as QC to remove spots that don't meet metrics. Includes among others:
1. Library size
2. # of expressed genes
3. Proportion of gene specific reads (mitochondrial in this example)
4. # of cells per spot (If available for dataset)

```{r}
hist(colData(spe1)$sum, breaks = 20) # Library Size, total sum of UMI counts per spot (included in sum column)

hist(colData(spe1)$detected, breaks = 20) # Number of expressed genes per spot (detected column)

hist(colData(spe1)$subsets_mito_percent, breaks = 20) # Proportion of mitochondrial genes per spot

# select QC threshold for library size
qc_lib_size <- colData(spe1)$sum < 5000
table(qc_lib_size)

colData(spe1)$qc_lib_size <- qc_lib_size

# check spatial pattern of discarded spots
plotQC(spe1, type = "spots", 
       discard = "qc_lib_size")

# # combined set of discarded spots
# discard <- qc_lib_size | qc_detected | qc_mito | qc_cell_count
# colData(spe)$discard <- discard

# # filter low-quality spots
# spe <- spe[, !colData(spe)$discard]

```
Log based normalization can be performed to scale data approximately and create clusters.
Is this necessary for filtered data?


``` {r}
set.seed(123) #quick clustering for pool-based size factors
qclus <- quickCluster(spe1)
# table(qclus)

spe1 <- computeSumFactors(spe1, cluster = qclus)
summary(sizeFactors(spe1))
hist(sizeFactors(spe1), breaks = 20)

spe1 <-  logNormCounts(spe1)
assayNames(spe1)
```
Feature selection based on highly variable genes (HVG) and spatially variable genes (SVG).
Should mitochondrial genes be removed? Are they are likely highly expressed and/or are they of biological interest?
SVG selection based on SPARK-X package, HVG based on scran package

``` {r}
# spe1 <- spe1[!is_mito, ]
# dim(spe1)

deCells <-  modelGeneVar(spe1)

fit <-  metadata(deCells)
plot(fit$mean, fit$var, 
     xlab = "mean of log-expression", ylab = "variance of log-expression")
curve(fit$trend(x), col = "dodgerblue", add = TRUE, lwd = 2)

top_hvgs <- getTopHVGs(deCells, prop = 0.1) #prop defines how many HVGS are wanted, 0.1 returns top 10% of genes
length(top_hvgs)

```
PCA and UMAP can be easily produced.
How do PCA/UMAP plots compare to the plots prpduced by loupe browser for 6week sample?
  They are different but I think different because they have 8 clusters, Compared to my 6. 


```{r}
set.seed(123)
spe1 <- runPCA(spe1, subset_row = top_hvgs)
spe1 <- runUMAP(spe1, dimred = "PCA")

# dim(reducedDim(spe, "PCA"))

colnames(reducedDim(spe1, "UMAP")) <- paste0("UMAP", 1:2)

plotDimRed(spe1, type = "PCA")
plotDimRed(spe1, type = "UMAP")

```

Produces clusters of HVGs using graph-based clustering based on the Walktrap method. Applies to top 50 PCs? (Meaning?)

Change K values to get number of clusters less than 8. Are less clusters better or are more clusters more accurate?
```{r}
set.seed(123)
k = 23
g <- buildSNNGraph(spe1, k = k, use.dimred = "PCA") #Graph based clustering
g_walk <- igraph::cluster_walktrap(g)
clus <- g_walk$membership
table(clus)

colLabels(spe1) <- factor(clus) #Store cluster labels in column 'label' in colData

plotSpots(spe1, annotate = "label",
          palette = "libd_layer_colors")

plotVisium(spe1, fill = "label", trans = "log", highlight = "in_tissue")

plotDimRed(spe1, type = "PCA",
           annotate = "label", palette = "libd_layer_colors")

plotDimRed(spe1, type = "UMAP",
           annotate = "label", palette = "libd_layer_colors")
```
Code to identify marker genes, can also be done using Seurat

``` {r}
rownames(spe1) <- rowData(spe1)$symbol

markers <- findMarkers(spe1, test = 'binom', direction = 'up') #Produces list of clusters equal to clusters from graph based
markers

interesting <- markers[[6]] #selection for which cluster
best_set <- interesting[interesting$Top <= 5, ]
logFCs <- getMarkerEffects(best_set)

pheatmap(logFCs, breaks = seq(-5, 5, length.out = 101))

top_genes <- head(rownames(interesting))
plotExpression(spe1, x = "label", features = top_genes)

```

